---
title: "Lab 05 -- Data Analysis"
author: "Melissa Juarez"
date: "March 28, 2023"
output:
  html_document: default
---

```{r include=FALSE}
library(here)
library(tidyverse)
library(janitor)
library(plyr)
library(dplyr)

## read data
wvs <- read.csv(here::here("data/WVS.csv"))
```

##### Q1: Run a multiple linear probability model.  Tell me how you think your independent variables will affect your dependent variable.  Interpret your results.  Were your expectations correct?  Why or why not?

```{r q1, message=FALSE, warning=FALSE}

```



##### Q2: Run a multiple (binary) logistic model.  (It can be the same as the above LPM or -- even better -- a new model.)  Tell me how you think your independent variables will affect your dependent variable.  Interpret your results in the logit scale.  Were your expectations correct?  Why or why not?

```{r q2}

```


##### Q3: Get odds ratios from your logit model in Question 2 and interpret some of them.  


```{r q3}

```


##### Q4: Get predicted probabilities from your logit model in Question 2 for some constellations of X values and interpret the results.  

```{r q4}

```


